\bf DOGE Days on Reddit: Decoding Public Sentiment in a Federal Shakeup

### 2025-04-07 code/ml_llm/R/assign_3.tex

---
title: "DOGE Days on Reddit: Decoding Public Sentiment in a Federal Shakeup"
format:
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
  jasa-html: default
date: last-modified
author:
  - name: Kevin Linares
    affiliations:
      - name: University of Maryland
  - name: Felix Baez-Santiago
  - name: Aria Lu
  - name: Gloria Zhou
    affiliations:
      - name: University of Michigan
abstract: |
   This study investigates public sentiment on Reddit towards the Department of Government Efficiency's (DOGE) federal workforce reduction. Analyzing 400 labeled Reddit comments, we employed supervised learning (KNN, Random Forest) and unsupervised Large Language Models (LLMs: Llama 3.2 3B, deepseek-r1:8b) to classify comment stance (favor, neutral, oppose). While LLMs showed higher precision for opposing comments, all models struggled with neutral and favoring stances. Agreement between the two LLMs was fair (Cohen's Kappa = 0.344), indicating divergent interpretations. Results suggest that accurately decoding nuanced public sentiment in this context remains challenging for both traditional and large language models.

keywords:
  - Reddit
  - Federal Government
  - DOGE
editor: 
  markdown: 
    wrap: sentence
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, comment = FALSE)

library(readxl)
library(viridis)
library(ggthemes)
library(tidytext)
library(knitr)
library(ConfusionTableR)
library(caret)
library(tidyverse)

# read in ML results from repo
results_folder <- "~/repos/SURV622_Assignment/outputs/ml_llm/"

ml_results <-  read.csv(str_c(
  results_folder, "results_long.csv")) |> 
  filter( data == "counts",
    set == "test",# keep only test results
    metric %in% c("precision", "recall", "f1-score"),
    class %in% c("favor", "neutral", "oppose"),
    !str_detect(model, "Bernoulli")
         ) |> 
  # remove duplicates
  distinct() |> 
  mutate(model = str_remove(model,  "^[^|]+\\.")) |> 
  tibble()

llm_results <- read_xlsx(str_c(
  results_folder, "reddit_comments_LLM_analysis.xlsx")) |> 
  mutate_at(vars(outcome, llama, deepseek), factor)
```

## Introduction {#sec-intro}

@@

## Methods {#sec-meth}

*Data*.
We collected Reddit comments from subreddits related to the topic of interest in early March of 2024.
This resulted in 12,553 comments which we preprocessed by removing weblinks, replace special characters (i.e., replaced "\@" with "at"), replaced numeric values with their spelling, and confirmed the absence of duplicate comments.
We then randomly selected 400 comments (without replacement) and assigned them to four graduate students for coding, categorizing each comment as favoring, opposing, or neutral towards DOGE's approach to federal workforce reductions.
Table 1 presents the breakdown in percent of our the comments we labelled and later use to train and evaluate our machine learning models.
A review of the comments revealed more negative opinionated statements and reactions.

```{r}
llm_results |> 
  count(outcome) |> 
  mutate(Percent = n/sum(n),
         Percent = round(Percent* 100)) |> 
  select(outcome, Percent) |> 
  kable(caption="Distribution of labeled Reddit comment data.")
```

@@

In addition to our supervised machine learning models, we also applied to our data two LLMs, llama 3.2.
3B and deepseek-r1:8b.
The llama model, developed by Meta AI released in February 2023, was designed for applications requiring low-latency inferencing with limited computational resources, which means that we were able to also deploy these models on a local machine with a GPU in addition to using the [Great Lakes HCP](https://its.umich.edu/advanced-research-computing/). This model also excels at text classification which is how we used it to classify the stance for each comment.
Deepseek, developed by Liang Wenfeng in July 2023 and based on the llama architecture, yet boosts efficiency improvements that appear to provide advances in reasoning, logic, while excelling at natural language processing.
We choose both of these LLMs as candidates to test against our supervised approaches because of their ability to apply locally and reputable performance on zero-shot classification tasks.
Both of these models received the same prompt and tasked which were developed by the research team to reflect the utility of classifying stance within our topic of interest.
Both LLMs resulted in not being able to classify five comments each while they only had one in common with the likely reason being that these comments did not contain enough useful text for a stance to be derive.

Prompt: "Is this comment in 'favor', 'neutral', or 'oppose' the reduction in federal workforce? Provide one word answer only!

Task: "You have assumed the role of a stakeholder that is presented with a reddit comment from a likely federal worker related to the current policies on reducing the federal workforce. Please determine the author's stance on this topic, and only provide the answer."

@@

## Results

@@

```{r}
#| fig-width: 7
#| fig-height: 6
#| 
# number of total posts
ml_results |> 
  # filter models needed
  filter(model %in% 
           c("RandomForestClassifier", "KNeighborsClassifier"),
         metric %in% c("precision", "recall")) |> 
  ggplot(aes(x=model, y=value, fill=class)) +
  geom_col(position="dodge") +  
  coord_flip() +
  facet_grid(~metric) +
  guides(fill=guide_legend(title="", reverse = TRUE)) +
  scale_fill_viridis_d(option="F", direction = -1,
                       alpha=.85, end =.70, begin=.15) +
  labs(
    title = "Figure 1. Recall and precision scores for random forest
    and KNN models.",
    x = "Models",
    y = "Performance Scores"
  ) +
  theme_hc()

```

@@

\

```{r}

 cm_metric_fun <- function(dat, model_name){

   subset_dat <- dat |> 
     # truth should go first, predicted 2nd
     select(outcome, all_of(model_name)) |> 
     rename(predictor = 2)

  cm_ob = multi_class_cm(subset_dat$outcome, 
                        subset_dat$predictor,
                        mode="everything")

  result = cm_ob$record_level_cm |> 
   tibble() |> 
   select(contains(c("precision", "recall", "f1"))) |> 
   pivot_longer(everything()) |> 
   separate(name, c("class", "metric")) |> 
   mutate(model = model_name) |> 
   rename(score=value)

  return(result)
 }

 llama_metrics <- cm_metric_fun(
   dat = llm_results, 
   model_name = "llama") # name of predicted column

 deepseek_metrics <- cm_metric_fun(
   dat = llm_results, 
   model_name = "deepseek" # name of predicted column
 )

 metric_results <- llama_metrics |> 
   add_row(deepseek_metrics)

```

```{r}
#| fig-width: 7
#| fig-height: 6
#| 
metric_results |> 
  filter(metric %in% c("Precision", "Recall")) |> 
  ggplot(aes(x=model, y=score, fill=class)) +
  geom_col(position="dodge") +  
  coord_flip() +
  facet_grid(~metric) +
  guides(fill=guide_legend(title="", reverse = TRUE)) +
  scale_fill_viridis_d(option="F", direction = -1,
                       alpha=.85, end =.70, begin=.15) +
  labs(
    title = "Figure 2. Recall and precision scores for LLMs.",
    x = "Models",
    y = "Performance Scores"
  ) +
  theme_hc()
```

Llama and DeepSeek demonstrated only fair agreement in stance classification (Cohen's Kappa = 0.344).
The contingency table, Table 2 below, reveals highest agreement for "oppose" (0.58), indicating some consistency in identifying strong negative sentiment.
However, agreement was substantially lower for "neutral" (0.11) and "favor" (0.02), highlighting divergent interpretations of more nuanced language.
Off-diagonal values further illustrate these discrepancies, suggesting fundamental differences in how the models process ambiguous cues and establish classification boundaries, particularly for less extreme stances.

```{r}
cm_results <- confusionMatrix(llm_results$llama, llm_results$deepseek)

coh_kappa <-  cm_results[[3]][[2]]

cm_results$table |> 
    kable(caption=str_c("Aggreement between LLMs, Cohens Kappa, ", 
                        round(coh_kappa, 2)))

```

## Conclusion

We explored public sentimenton Reddit regarding the DOGE federal workforce reduction using supervised learning (KNN, Random Forest) and unsupervised LLMs (Llama 3.2 3B, deepseek-r1:8b) for stance classification.
LLMs showed higher precision for "oppose" comments, suggesting better identification of negative sentiment, but lower recall, indicating missed opposing comments.
Supervised models performed consistently but generally poorly across all stances, and likely due to the limited feature set (Reddit metadata) as well as the small sample of comments.

The consistent difficulty in classifying "neutral" and "favor" stances across all models likely stems from class imbalance (more "oppose" comments), the inherent ambiguity of neutral language, and potentially more nuanced expressions of favor.
The zero-shot approach for LLMs might also limit their performance on this specific domain.
The few unclassifiable comments highlight the inherent challenges in stance detection within informal online text.

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={DOGE Days on Reddit: Decoding Public Sentiment in a Federal Shakeup},
  pdfauthor={Kevin Linares; Felix Baez-Santiago; Aria Lu; Gloria Zhou},
  pdfkeywords={Reddit, Federal Government, DOGE},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{April 7, 2025}
\title{\bf DOGE Days on Reddit: Decoding Public Sentiment in a Federal
Shakeup}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
This study investigates public sentiment on Reddit towards the
Department of Government Efficiency's (DOGE) federal workforce
reduction. Analyzing 400 labeled Reddit comments, we employed supervised
learning (KNN, Random Forest) and unsupervised Large Language Models
(LLMs: Llama 3.2 3B, deepseek-r1:8b) to classify comment stance (favor,
neutral, oppose). While LLMs showed higher precision for opposing
comments, all models struggled with neutral and favoring stances.
Agreement between the two LLMs was fair (Cohen's Kappa = 0.344),
indicating divergent interpretations. Results suggest that accurately
decoding nuanced public sentiment in this context remains challenging
for both traditional and large language models.
\end{abstract}

\noindent%
{\it Keywords:} Reddit, Federal Government, DOGE
\vfill

\newpage
\spacingset{1.9} % DON'T change the spacing!

\section{Introduction}\label{sec-intro}

@@

\section{Methods}\label{sec-meth}

\emph{Data}. We collected Reddit comments from subreddits related to the
topic of interest in early March of 2024. This resulted in 12,553
comments which we preprocessed by removing weblinks, replace special
characters (i.e., replaced ``@'' with ``at''), replaced numeric values
with their spelling, and confirmed the absence of duplicate comments. We
then randomly selected 400 comments (without replacement) and assigned
them to four graduate students for coding, categorizing each comment as
favoring, opposing, or neutral towards DOGE's approach to federal
workforce reductions. Table 1 presents the breakdown in percent of our
the comments we labelled and later use to train and evaluate our machine
learning models. A review of the comments revealed more negative
opinionated statements and reactions.

\begin{longtable}[]{@{}lr@{}}
\caption{Distribution of labeled Reddit comment data.}\tabularnewline
\toprule\noalign{}
outcome & Percent \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
outcome & Percent \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 28 \\
neutral & 18 \\
oppose & 54 \\
\end{longtable}

@@

In addition to our supervised machine learning models, we also applied
to our data two LLMs, llama 3.2. 3B and deepseek-r1:8b. The llama model,
developed by Meta AI released in February 2023, was designed for
applications requiring low-latency inferencing with limited
computational resources, which means that we were able to also deploy
these models on a local machine with a GPU in addition to using the
\href{https://its.umich.edu/advanced-research-computing/}{Great Lakes
HCP}. This model also excels at text classification which is how we used
it to classify the stance for each comment. Deepseek, developed by Liang
Wenfeng in July 2023 and based on the llama architecture, yet boosts
efficiency improvements that appear to provide advances in reasoning,
logic, while excelling at natural language processing. We choose both of
these LLMs as candidates to test against our supervised approaches
because of their ability to apply locally and reputable performance on
zero-shot classification tasks. Both of these models received the same
prompt and tasked which were developed by the research team to reflect
the utility of classifying stance within our topic of interest. Both
LLMs resulted in not being able to classify five comments each while
they only had one in common with the likely reason being that these
comments did not contain enough useful text for a stance to be derive.

Prompt: ``Is this comment in `favor', `neutral', or `oppose' the
reduction in federal workforce? Provide one word answer only!

Task: ``You have assumed the role of a stakeholder that is presented
with a reddit comment from a likely federal worker related to the
current policies on reducing the federal workforce. Please determine the
author's stance on this topic, and only provide the answer.''

@@

\section{Results}\label{results}

@@

\includegraphics{assign_3_files/figure-pdf/unnamed-chunk-3-1.pdf}

@@

\hfill\break

\includegraphics{assign_3_files/figure-pdf/unnamed-chunk-5-1.pdf}

Llama and DeepSeek demonstrated only fair agreement in stance
classification (Cohen's Kappa = 0.344). The contingency table, Table 2
below, reveals highest agreement for ``oppose'' (0.58), indicating some
consistency in identifying strong negative sentiment. However, agreement
was substantially lower for ``neutral'' (0.11) and ``favor'' (0.02),
highlighting divergent interpretations of more nuanced language.
Off-diagonal values further illustrate these discrepancies, suggesting
fundamental differences in how the models process ambiguous cues and
establish classification boundaries, particularly for less extreme
stances.

\begin{longtable}[]{@{}lrrr@{}}
\caption{Aggreement between LLMs, Cohens Kappa, 0.34}\tabularnewline
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 6 & 7 & 1 \\
neutral & 3 & 42 & 14 \\
oppose & 29 & 61 & 228 \\
\end{longtable}

\section{Conclusion}\label{conclusion}

We explored public sentimenton Reddit regarding the DOGE federal
workforce reduction using supervised learning (KNN, Random Forest) and
unsupervised LLMs (Llama 3.2 3B, deepseek-r1:8b) for stance
classification. LLMs showed higher precision for ``oppose'' comments,
suggesting better identification of negative sentiment, but lower
recall, indicating missed opposing comments. Supervised models performed
consistently but generally poorly across all stances, and likely due to
the limited feature set (Reddit metadata) as well as the small sample of
comments.

The consistent difficulty in classifying ``neutral'' and ``favor''
stances across all models likely stems from class imbalance (more
``oppose'' comments), the inherent ambiguity of neutral language, and
potentially more nuanced expressions of favor. The zero-shot approach
for LLMs might also limit their performance on this specific domain. The
few unclassifiable comments highlight the inherent challenges in stance
detection within informal online text.

@@

\end{document}
